Alone Together: Compositional Reasoning and Inference for Weak Isolation

### 2017-11-08 motivation.tex
CCS and other things.

Because the aforementioned reasoning is reasonably simple to perform
manually, verifying the soundness of TPC-C's consistency conditions
would appear to be feasible.  Serializability aids the tractability of
verification by preventing any interference among concurrently
executing transactions while the \C{new\_order} transaction executes,
essentially yielding serial behaviors.  Under weak
isolation\footnote{Weak isolation does not violate atomicity as long
  as the witnessed effects are those of committed transactions},
however, interferences of various kinds are permitted, leading to
executions superficially similar to executions permitted by concurrent
(racy) programs~\cite{GHE15,HPQ+15}. 
\begin{wrapfigure}{l}{.4\textwidth}
\includegraphics[scale=0.45]{Figures/motiv-eg-1-b}
\caption{\small An RC execution involving two instances ($T_1$ and
  $T_2$) of the \C{new\_order} transaction depicted in
  Fig.~\ref{fig:new_order_code}. 
  Both instances read the \C{d\_id} \C{District} record concurrently,
  because neither transaction is committed when the reads are
  executed.  The subsequent operations are effectively sequentialized,
  since $T_2$ commits before $T_1$. Nonetheless, both transactions read the same value for
  \C{d\_next\_o\_id} resulting in them  adding \C{Order} records
  with the same ids, which in turn triggers a violation of TPC-C's
  consistency condition.}
\label{fig:new_order_exec}
\end{wrapfigure}
To illustrate, consider the
behavior of the \C{new\_order} transaction when executed under a
\emph{Read Committed} (RC) isolation level, the default isolation
level in 8 of the 18 databases studied in~\cite{bailishotos}.
An executing RC transaction is isolated from \emph{dirty writes},
i.e., writes of uncommitted transactions, but is allowed to witness
the writes of concurrent transactions as soon as they are committed.
Thus, with two concurrent instances of the \C{new\_order} transaction
(call them $T_1$ and $T_2$), both concurrently placing new orders for
different customers under the same district (\C{d\_id}), RC isolation
allows the execution shown in Fig.~\ref{fig:new_order_exec}.

>>>>

Because the aforementioned reasoning is reasonably simple to perform
manually, verifying the soundness of TPC-C's consistency conditions
would appear to be feasible.  Serializability aids the tractability of
verification by preventing any interference among concurrently
executing transactions while the \C{new\_order} transaction executes,
essentially yielding serial behaviors.  
\begin{wrapfigure}{l}{.4\textwidth}
\includegraphics[scale=0.45]{Figures/motiv-eg-1-b}
\caption{\small An RC execution involving two instances ($T_1$ and
  $T_2$) of the \C{new\_order} transaction depicted in
  Fig.~\ref{fig:new_order_code}. 
  Both instances read the \C{d\_id} \C{District} record concurrently,
  because neither transaction is committed when the reads are
  executed.  The subsequent operations are effectively sequentialized,
  since $T_2$ commits before $T_1$. Nonetheless, both transactions read the same value for
  \C{d\_next\_o\_id} resulting in them  adding \C{Order} records
  with the same ids, which in turn triggers a violation of TPC-C's
  consistency condition.}
\label{fig:new_order_exec}
\vspace*{-10pt}
\end{wrapfigure}
Under weak isolation\footnote{Weak isolation does not violate
  atomicity as long as the witnessed effects are those of committed 
transactions}, however, interferences of various kinds are permitted,
leading to executions superficially similar to executions permitted by
concurrent (racy) programs~\cite{GHE15,HPQ+15}.  To illustrate,
consider the behavior of the \C{new\_order} transaction when executed
under a \emph{Read Committed} (RC) isolation level, the default
isolation level in 8 of the 18 databases studied
in~\cite{bailishotos}.  An executing RC transaction is isolated from
\emph{dirty writes}, i.e., writes of uncommitted transactions, but is
allowed to witness the writes of concurrent transactions as soon as
they are committed.  Thus, with two concurrent instances of the
\C{new\_order} transaction (call them $T_1$ and $T_2$), both
concurrently placing new orders for different customers under the same
district (\C{d\_id}), RC isolation allows the execution shown in
Fig.~\ref{fig:new_order_exec}.
